---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# pxmake

<!-- badges: start -->
![R-CMD-check](https://github.com/StatisticsGreenland/pxmake/actions/workflows/R-CMD-check.yml/badge.svg)
<!-- badges: end -->

pxmake is an R package for creating and modifying px files.

With pxmake you can:

-   Import a PC-AXIS px file, modify it, and save it as a new px file.
-   Save a px file as an Excel workbook.
-   Do complex modifications to a px file, like adding total levels to a variable.

## Installation

Install the latest version by running:

``` r
# install.packages('devtools')
devtools::install_github('StatisticsGreenland/pxmake')
```

pxmake is not available on CRAN.

## How to use

Use `px()` and `pxsave()` to read/write a px file to a px object.

```{r include=FALSE, messages = FALSE}
library(pxmake)
set.seed(1)
```


``` r
library(pxmake)

# Read px file into a px object
x <- px(input = "example.px")

class(x)
#> [1] "px"

# Save px object as a new px file
pxsave(x, path = "example2.px")
```
To modify a px object, use pxmake's *modifying functions*. 

In general, modifying functions are named after the keyword they modify. It's possible to chain multiple modifying functions together in tidyverse style by using the pipe operator `%>%`.

```{r}
library(magrittr) # import pipe

# Create px object from data frame
x <- px(data.frame(year = as.character(rep(2021:2023, each = 3)), 
                   group = c('a', 'b', 'c'), 
                   value = runif(9)
                   )
        ) 

head(x$data, 4)

x %>% 
  timeval("year") %>%  # Set year as TIMEVAL
  heading("year") %>%  # Set year as HEADING
  stub("group") %>%    # Set group as STUB
  decimals("2") %>%    # Set DECIMALS to 2
  pxsave("example.px") # Save as px file
```

```{r echo = FALSE}
unlink("example.px")
```


### Modifying functions
Currently the following keywords have a modifying function in pxmake:

```{r, include = FALSE}
modifying_functions <-
  readLines('NAMESPACE') %>%
  stringr::str_match("S3method\\((?<function>.*),px\\)") %>%
  as.data.frame() %>%
  tidyr::drop_na() %>% 
  dplyr::pull("function")

status <- 
  readLines(file.path('R', 'keywords.R')) %>% 
  tail(-8) %>% 
  stringr::str_match("\\S(?<Keyword>[^,]*)\",") %>% 
  as.data.frame() %>% 
  tidyr::drop_na() %>% 
  dplyr::select(Keyword) %>% 
  dplyr::mutate(`Function name` = gsub("-", "_", tolower(Keyword)),
                implemented = `Function name` %in% modifying_functions
                )
```

```{r, echo = FALSE}
status %>% 
  dplyr::filter(implemented) %>%
  dplyr::select(-implemented) %>% 
  print(row.names = FALSE)
```
In addition to the above, the following modifying functions are available:

```{r, echo = FALSE}
dplyr::tibble(`Function name` = setdiff(modifying_functions, status[, "Function name"])) %>% 
  as.data.frame() %>% 
  print(row.names = FALSE)
```

See the help page for each modifying function for more information.

<details>
<summary>Keywords without modifying functions</summary>
These keywords currently doesn't have a modifying function, but most of them can be implemented.
```{r, echo = FALSE}
status %>% 
  dplyr::filter(!implemented) %>%
  dplyr::select(-implemented) %>% 
  print(row.names = FALSE)
```
</details>

## Need help?
If you have any questions or need help, feel free to [open an issue](https://github.com/StatisticsGreenland/pxmake/issues/new) on GitHub, or contact [Lars and Johan](https://github.com/StatisticsGreenland/pxmake/graphs/contributors) via email.

## For developers

### PX specification

See the [PX-file format specification on Statistics Swedens homepage](https://www.scb.se/globalassets/vara-tjanster/px-programmen/px-file_format_specification_2013.pdf).

### PxJob

Some tests cases uses [PxJob](https://www.stat.fi/tup/tilastotietokannat/px-tuoteperhe_en.html). Install [pxjob64Win](https://github.com/StatisticsGreenland/pxjob64Win) to be able ro run these tests. This only works on Windows.

### How to create a new release

1.  Checkout 'main' branch.
2.  Run `usethis::use_version('major'/'minor'/'patch')`. Answer 'No' to the prompt 'Is it ok to commit them?'.
3.  Update NEWS.md with all changes since last release.
4.  Commit changes with message 'Increment version number to X.Y.Z'.
5.  Run `git tag vX.Y.Z`.
6.  Run `git push`.
7.  Run `git push --tags`.
8.  Run `devtools::use_github_release()`.
9.  Review relase description. If sattified, click the 'edit' pen in the upper right corner.
10. Click 'Publish release'.
11. Run `usethis::use_version('dev')`. Answer 'Yes' to the prompt 'Is it ok to commit them?'.
12. Run `git push`.
