---
title: "Languages"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Languages}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The px file format supports having multiple langauges in one file. Let's start by creating a minimal example of a px object, based on the built in data set `population_gl`.

```{r setup}
library(pxmake)

x <- px(population_gl)
```
Currently, `x` has neither a main language nor any additional languages.

```{r}
px_language(x)  # Main language
px_languages(x) # All languages
```

To set the main language, use `px_language()`, and add the [ISO 639](https://en.wikipedia.org/wiki/List_of_ISO_639_language_codes) language code.

```{r}
x1 <- px_language(x, "kl")
```

This adds the keyword 'LANGUAGE=kl' if the px file is saved, but otherwise there are no changes. 

To set multiple languages, use `px_languages()` with a list of language codes.

```{r}
x2 <- px_languages(x1, c("da", "kl"))
```

This doubles the number of values for all language dependent keywords. For example CONTACT, can be set simultaneously for both languages:

```{r}
x3 <- px_contact(x2, "Johan Ejstrud")
```

or it can be set with distinct values for the two languages:

```{r}
x4 <- px_contact(x3, dplyr::tribble(~language, ~value,
                                    "kl", "Lars Pedersen",
                                    "da", "Johan Ejstrud"
                                    )
                 )
```

In general, changing a keyword for multiple languages, requires a data frame where the column 'language' has been added. See the help pages for each functions to see the exact details on which columns are needed in the data frame.
