---
title: "my first px file"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{my-first-px-file}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(pxmake)
```
## Getting started

This code turns a simple dataframe into a **px**-file with just enough metadata to open with **pxwin**

Create a sample dataframe, **df**

```{r sample_data, message=FALSE}

library(tidyverse)

df <-
  expand_grid(sex = c("Female", "Male"),
              age = c("0-15", "16-25", "26-50"),
              time = c("2021", "2022", "2023")
              ) %>%
  mutate(value = round(rnorm(nrow(.), mean = 20, sd = 5), 0)) %>% 
  as_tibble()

df %>% print(n=5)

```

use the **pxmake** library to convert the dataframe to **.rds**, **.px** and **.xlsx**, like this:

```{r}

# devtools::install_github('StatisticsGreenland/pxmake')
library(pxmake)

# from dataframe to rds
rds <- metamake(df)

# MATRIX - used also for filename in this example
px_matrix <- "BEXFIRST"

# from dataframe to xlsx
metamake(df, paste0(px_matrix,".xlsx"))

# from rds to px
pxmake(rds,paste0(px_matrix,".px"))

# view(rds$metadata)
# view(rds$data)
```

The sample dataframe can be edited in Excel or R. Metadata is still too vaguely defined for **pxwin** to show the file, but **pxedit** does (of course).

For **pxwin** to show the **px**-file, these keywords cannot be blank:\
MATRIX, CONTENTS, UNITS, SUBJECT-CODE, SUBJECT-AREA

So update the keywords (can alternatively be edited in Excel):

```{r}
# Update keyword(s) 

rds$metadata <-
  rds$metadata %>%
  mutate(
    value = ifelse(keyword == "MATRIX",        px_matrix,           value),
    value = ifelse(keyword == "CONTENTS",     "Sample persons",     value),
    value = ifelse(keyword == "UNITS",        "Persons",            value),
    value = ifelse(keyword == "SUBJECT-CODE", "BE",                 value),
    value = ifelse(keyword == "SUBJECT-AREA", "Population",         value)
    )

# save metadata changes
metamake(rds, paste0(px_matrix,".xlsx"))

# convert to px-file
pxmake(input = paste0(px_matrix,".xlsx"), 
       out_path = paste0(px_matrix,".px"))


```

and the **px**-file will show in **pxwin**

Additional keywords can be added, like this:

```{r}

# add keywords
rds$metadata <-
  rds$metadata %>%
  bind_rows(tribble(~keyword, ~language,~variable,~value,
                  "NOTEX","en","age",list("Rough age groups only")))

# save metadata changes
metamake(rds, paste0(px_matrix,".xlsx"))

# convert to px-file
pxmake(input = paste0(px_matrix,".xlsx"), 
       out_path = paste0(px_matrix,".px"))


```

as can totals. First add metadata, and use the pxmake, add_totals setting

```{r}

# add metadata for totals 
rds$metadata <-
  rds$metadata %>%
  filter(keyword != "ELIMINATION") %>%
  bind_rows(tibble(keyword  = "ELIMINATION",
                   language = "en",
                   variable = c("age", "sex"),
                   value    = list("Total")
                   ))

# save metadata changes
metamake(rds, paste0(px_matrix,".xlsx"))

# convert to px-file and calculate totals
pxmake(input = paste0(px_matrix,".xlsx"), 
       out_path = paste0(px_matrix,".px"),
       add_totals = c("age","sex"))
```
