---
title: "Statistics Greenland"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Statistics Greenland}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## Minimal px template

Below is a minimal template for tables meant to be presented in Statbank Greenland. 

```{r, warning = FALSE, message = FALSE}
library(pxmake)
library(dplyr)
library(stringr)
```

```{r,}
creation_date <- "20240301 09:00"
last_updated <- "20240711 09:00"
next_update   <- "20250301 09:00"
year <- "2024"

subject <-
  tribble(~language, ~value,
          "en", "Population", 
          "da", "Befolkning", 
          "kl", "Innuttaasut")

description <- 
  subject %>% 
  mutate(value = str_glue("{value} 1999-{year} <em>[BEEP24MORT]</em>"))

x <-
  population_gl %>%
  pxmake::px() %>%
  px_codepage("utf-8") %>%
  px_language("en") %>%
  px_languages(c("en", "da", "kl")) %>%
  px_decimals("15") %>%
  px_showdecimals("0") %>%
  px_matrix("POP") %>%
  px_subject_code("BE") %>%
  px_subject_area(subject) %>%
  px_contents(subject) %>%
  px_stub(c("gender", "age")) %>%
  px_heading("year") %>%
  px_timeval("year") %>%
  px_description(description) %>%
  px_title(subject) %>%
  px_units(tribble(~language, ~value,
                   "en", "Persons", 
                   "da", "Personer", 
                   "kl", "Inuit")) %>%
  px_domain(tribble(~`variable-code`, ~language,    ~domain,
                              "age" ,      "en",    "VPAge",
                              "age" ,      "da",  "VPAlder",
                              "age" ,      "kl",  "VPUkiut")) %>%
  px_update_frequency("Annually") %>%
  px_creation_date(creation_date) %>%
  px_last_updated(last_updated) %>%
  px_next_update(next_update) %>%
  px_link("www.stat.gl/bee202401/m1") %>%
  px_contact("Lars Pedersen, LARP at STAT.gl") %>%
  px_source(tribble(~language, ~value,
            "en",  "Statistics Greenland",
            "da",  "Grønlands Statistik",
            "kl",  "Kalaallit Nunaanni Naatsorsueqqissaartarfik"))

```

## Subjects

px files are published in Statbank Greenland and organized into subjects/topics. The api (pxweb) exposes this information for each supported language in  https://bank.stat.gl/api/v1/{language}/Greenland.
  

### Subjects by ChatGPT

ChatGPT can be used to generate R code. In this example, api endpoints adresses to subject (codes and texts) are provided and ChatGPT i asked to:  

Please combine data from  
https://bank.stat.gl/api/v1/en/Greenland,  
https://bank.stat.gl/api/v1/da/Greenland and  
https://bank.stat.gl/api/v1/kl/Greenland  
into one dataframe in r and add column with language, in tidyverse, fetch as function

This is returned:  

```{r, eval = FALSE}
by ChatGPT

# Load the necessary packages
library(httr)
library(jsonlite)
library(tidyverse)

# Function to fetch data from API and add language column
fetch_data <- function(url, language) {
  response <- GET(url)
  data <- content(response, as = "text")
  json_data <- fromJSON(data, flatten = TRUE) %>% as_tibble()
  json_data <- json_data %>% mutate(language = language)
  return(json_data)
}

# URLs for the APIs
url_en <- "https://bank.stat.gl/api/v1/en/Greenland"
url_da <- "https://bank.stat.gl/api/v1/da/Greenland"
url_kl <- "https://bank.stat.gl/api/v1/kl/Greenland"

# Fetch data from each API
df_en <- fetch_data(url_en, 'English')
df_da <- fetch_data(url_da, 'Danish')
df_kl <- fetch_data(url_kl, 'Greenlandic')

# Combine the dataframes
combined_df <- bind_rows(df_en, df_da, df_kl)

# Clean and inspect the combined dataframe
# View the first few rows of the combined dataframe
head(combined_df)

# Check the structure of the combined dataframe
glimpse(combined_df)

```

